---
title: "Project 1 - Proposal and Data Assembly"
author: "Andrew Kerr"
date: "2023-10-15"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
options(digits=4)
```

```{r, message=FALSE, include=FALSE, warning=FALSE}
library(tidyverse)
library(RPostgres)
library(RPostgreSQL)
library(kableExtra)
load(url("http://www.rossmanchance.com/iscam3/ISCAM.RData"))
```

```{r, include=FALSE}
# Replace these values with your Supabase database details
host <- "db.bnnpxoerfnicvlkamcdv.supabase.co"
dbname <- "postgres"
user <- "postgres"
password <- "0anadhrVeMVHTXxq"

# Establish a connection to the Supabase database
con <- dbConnect(RPostgres::Postgres(), 
                 host = host, 
                 dbname = dbname, 
                 user = user, 
                 password = password)

# match_history <- dbGetQuery(con, "SELECT * FROM match_history")
# match_info <- dbGetQuery(con, "SELECT * FROM match_info")
# players <- dbGetQuery(con, "SELECT * FROM players")
# rank_info <- dbGetQuery(con, "SELECT * FROM rank_info")

q = "
SELECT ri.tier, mh.match_id, p.name, mh.game_length, mh.tft_game_type, mi.placement, mi.level, 
mi.last_round, mi.gold_left, mi.augment_1, mi.augment_2, mi.augment_3, 
mi.traits, mi.units

FROM match_history as mh

JOIN match_info as mi ON mh.match_id = mi.match_id
JOIN rank_info as ri ON mi.puuid = ri.puuid
JOIN players as p ON ri.puuid = p.puuid

WHERE mh.tft_game_type = 'RANKED' AND mh.tft_set_core_name = 'TFTSet9_2'
"
df_TFT = dbGetQuery(con, q)


# Close connection to Supabase database
dbDisconnect(con)
```

# Research Question

One of my favorite games to play is Team Fight Tactics (TFT), a real-time 
strategy game centered around managing gold, augments (game modifiers different 
for each player), units, and traits (each unit has 1-3 traits), where 8 players 
fight it out to be the last one standing. 

One method of measuring the skill of a player is by looking at their tier. 
The tiers in TFT go as follows (from lowest to highest):

  + Iron
  + Bronze
  + Silver
  + Gold
  + Platinum
  + Diamond
  + Masters
  + Grandmasters
  + Challenger

For example, a Silver player is ranked lower than a Gold player.

I am interested in whether you can determine a players tier based on the results
of a match of TFT. In other words,

  + Are there distinct play styles or mannerism associated with different tiers 
  in TFT?

I have collected a random sample of matches from different tiers of ranked TFT.
The variables I have decided to include are:

  + placement: a measure of how well a player performed in a match compared to 
  other players in the match
    - I may create subgroups for each placement 
  + level: a measure of how well strong a player was in a match
  + last_round: what round a player is eliminated from a match
  + gold_left: amount of gold a player had when the were eliminated from a match
  + augment_1-3: game modifier selected by a player during the match
  + traits: active traits a player had when they were eliminated
  + units: units a player was using when they were eliminated
  + game_length: how long (in seconds) a match lasted
  + tier: what tier the player currently is

# Data

Over the summer I worked on using the Riot API (company that owns TFT) to 
collect my match history. I have re-purposed the code to now collect match
history for any tier of player. I provide the tier I desire, and my program 
acquires 205 randomly selected players from the tier and adds their 5 most 
recently played matches to my database. For each match I collect information
on all 8 players and how they performed in the match. Below is an observation 
from a data set created from a combination of my SQL tables (traits and units 
will need to be re-structured for this project):

```{r}
head(df_TFT, 1)
```

# Multileveled Structure

I plan to talk to you more about this, but here is what I am considering:

  + Observational units: individual player outcomes in a match
  + Level 1: player (8 players from each match)
  + Level 2: match (10k+ matches, I am currently collecting them overnight, there was a bug in my code :( ) 

# Variable Chart

```{r, include=FALSE}
tbl <- data.frame(Name = c("Tier", "placement", "level", "last round", "gold left", "legend augment count", "game length", "avg unit level", "traits", "avg level"),
                  VariableRole = c("Response", "Level 1 Predictor", "Level 1 Predictor", "Level 1 Predictor", "Level 1 Predictor", "Level 1 Predictor", "Level 2 Predicotr", "Level 1 Predictor", "Level 1 Predictor", "Level 2 Predictor"),
                  Type = c("Categorical", "Categorical", "Categorical", "Categorical", "Quantitative", "Categorical", "Quantitative", "Quantitative", "Categorical", "Quantitative"),
                  Values = c("Iron, Bronze, Silver, Gold, Platinum, Diamond, Masters, Grandmasters, Challenger", "1 to 8", "1 to 10", "1 to 50", "0 to 200", "how many legend augments were chosen => 0 to 3", "0 to 3600 (seconds)", "each unit can be between level 1 and 3 => 1 to 3", "list of active traits (will dummify; total of 28)", "mean level of a match"))
```

```{r}
kable(tbl, format = "simple")
```
