---
title: "Project 1 - Exploratory Data Analysis"
author: "Andrew Kerr"
date: "2023-10-29"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
options(digits=4)
```

```{r, message=FALSE, include=FALSE, warning=FALSE}
library(tidyverse)
library(jsonlite)
library(RPostgres)
library(RPostgreSQL)
load(url("http://www.rossmanchance.com/iscam3/ISCAM.RData"))
```

```{r, include=FALSE}
# Replace these values with your Supabase database details
host <- "db.bnnpxoerfnicvlkamcdv.supabase.co"
dbname <- "postgres"
user <- "postgres"
password <- "0anadhrVeMVHTXxq"
```

```{r}
# Establish a connection to the Supabase database
con <- dbConnect(RPostgres::Postgres(), 
                 host = host, 
                 dbname = dbname, 
                 user = user, 
                 password = password)

# match_history <- dbGetQuery(con, "SELECT * FROM match_history LIMIT 10000")
# match_info <- dbGetQuery(con, "SELECT * FROM match_info LIMIT 10000")
# players <- dbGetQuery(con, "SELECT * FROM players")
# rank_info <- dbGetQuery(con, "SELECT * FROM rank_info")

q = "
SELECT ri.tier, mh.match_id, p.name, mh.game_length, mi.placement, mi.level,
mi.last_round, mi.gold_left, mi.augment_1, mi.augment_2, mi.augment_3, mi.units

FROM match_history AS mh

JOIN match_info AS mi ON mh.match_id = mi.match_id
JOIN rank_info AS ri ON mi.puuid = ri.puuid
JOIN players AS p ON ri.puuid = p.puuid

WHERE mh.tft_game_type = 'RANKED'
AND mh.tft_set_core_name = 'TFTSet9_2'
"
df_TFT = dbGetQuery(con, q)

# Close connection to Supabase database
dbDisconnect(con)
```


# Data Collection and Cleaning

Using my program, I provide the tier I desire, and my program 
acquires 205 randomly selected players from the tier and adds their 5 most 
recently played matches to my database. For each match I collect information
on all 8 players and how they performed in the match. For the highest tier 
(Challenger), I was able to collect the 5 most recent matches of all players in 
this tier (n = 250).

The variables I have decided to include are:

  + placement: a measure of how well a player performed in a match compared to 
  other players in the match
  + level: a measure of how well strong a player was in a match
  + last_round: what round a player is eliminated from a match
  + gold_left: amount of gold a player had when the were eliminated from a match
  + augment_1-3: game modifier selected by a player during the match
  + units: units a player was using when they were eliminated
  + game_length: how long (in seconds) a match lasted
  + tier: what tier the player currently is

## Augments

To start I need to explain a bit more about the augment variables. All
players select 3 augments throughout each match, and there are two types of 
augments: normal augments and legend augments. Legend augments are tailored 
towards different play styles, and are chosen by a player prior to starting a 
match. Riot's API provides us the name of each augment a player chooses, unless 
it was a legend augment, in which case the API returns NA. However, there is a 
small chance that no augments are offered in a given match, and in this case the
API returns NA for all augments for all players in the match.

A new feature I have decided to create is the number of legend augments a player
chose during their match. Since I can not distinguish legend augments from a no 
augment match, I have dropped matches where all players have NA for all three
augments (this signifies that either everyone chose a legend augment 3 times, or
no augments were offered that match, the former being more unlikely than the 
latter; -63 observations).

```{r}
remove_matches <- df_TFT %>%
  group_by(match_id) %>%
  filter(is.na(augment_1), is.na(augment_2), is.na(augment_3)) %>%
  count() %>%
  filter(n == 8) %>%
  select(match_id)

# remove_matches %>% nrow()

df_TFT_clean <- df_TFT %>%
  filter(!(match_id %in% remove_matches)) %>%
  group_by(match_id, name) %>%
  mutate(n_legend = sum(
    if_else(last_round >= 5, is.na(augment_1), 0)
    + if_else(last_round >= 13, is.na(augment_2), 0)
    + if_else(last_round >= 20, is.na(augment_3), 0)
    ))
```

## Unranked Players

Below are the number of match results for each tier. Since we are not interested
in Unranked players, we will drop these match results from the data set (-1337 
observations).

```{r}
df_TFT_clean %>%
  group_by(tier) %>%
  select(tier, match_id) %>%
  count()

df_TFT_clean <- df_TFT_clean %>%
  filter(tier != 'UNRANKED')
```

## Units

Each unit a player has on their board has a cost of 1, 2, 3, 4, or 5 and can be 
level 1, 2, or 3. If you have 3 of the same level 1 units, it becomes a level 2,
and if you have 3 of the same level 2 units it becomes level 3 (e.g. the cost
of a level 2, 3 cost, unit is 6 gold). Two ways to measure how strong a players
board is is by looking at average unit level and total board cost. The following 
aggregates the units feature (currently a json) to be the average unit level and
total board cost the player ended the match with.

Additionally, since there is great variability in amount of resources (gold, 
free units, etc.) given to players from match to match, I have created features 
for average unit level and average board cost of all players in a match (level 2 
features). 

```{r}
df_units <- lapply(df_TFT_clean$units, fromJSON)

avg_units <- sapply(df_units, function(units) {
  if (length(units) > 0) {
    return(mean(units$tier))
  } 
  return(0)
})

board_costs <- sapply(df_units, function(units) {
  if (length(units) > 0) {
    cost <- units %>%
      filter(character_id != 'TFT9_HeimerdingerTurret') %>%
      # The rarity (cost) is off from the API results...
      mutate(costs = ifelse(rarity < 3, rarity + 1, ifelse(rarity == 6, rarity - 1, rarity)) * tier)
    return(sum(cost$costs))
  } 
  return(0)
})

df_TFT_clean$avg_unit_level <- avg_units
df_TFT_clean$board_cost <- board_costs

df_TFT_clean <- df_TFT_clean %>% select(-units)
```

## Independence

Since, within each match, knowing the results of one player can affect the 
probability of seeing certain results of another player (e.g. placement), I
have decided to randomly sample from my random sample. I will only keep 3 random
players per match.

```{r}
set.seed(43)

df_TFT_clean_sample <- df_TFT_clean %>%
  group_by(match_id) %>%
  slice_sample(n = 4, replace = FALSE)

df_TFT_clean_sample %>%
  group_by(tier) %>%
  select(tier, match_id) %>%
  count()

nrow(df_TFT_clean_sample)
```

```{r}
df_TFT_clean_sample <- df_TFT_clean_sample %>%
  select(-name) %>%
  mutate(
    game_length = as.integer(game_length),
    level = as_factor(level),
    placement = as_factor(placement),
    n_legend = as_factor(n_legend),
    match_id = as_factor(match_id),
    tier = as_factor(tier)
  )

df_TFT_clean_sample$tier <- factor(df_TFT_clean_sample$tier, levels=c('IRON', 'BRONZE', 'SILVER', 'GOLD', 'PLATINUM', 'DIAMOND', 'MASTER', 'GRANDMASTER', 'CHALLENGER'))
df_TFT_clean_sample$level <- factor(df_TFT_clean_sample$level, levels=c('1', '2', '3', '4', '5', '6', '7', '8', '9', '10'))
df_TFT_clean_sample$placement <- factor(df_TFT_clean_sample$placement, levels=c('1', '2', '3', '4', '5', '6', '7', '8'))
df_TFT_clean_sample$n_legend <- factor(df_TFT_clean_sample$n_legend, levels=c('0', '1', '2', '3'))
```


# Exploration

```{r}
plot(df_TFT_clean_sample[c(1, 3, 4, 5, 6)])
```

```{r}
plot(df_TFT_clean_sample[c(1, 7, 11, 12, 13)])
```

Based on the above scatter plots, it appears that there are a few outliers with
a 0 gold board cost and an average unit level of 0. This means that the players 
board was empty when the game ended for them. This should never realistically 
happen, and would only happen if a player has given up and does so on purpose. 
Thus, I will remove these outliers from the data.

```{r}
df_TFT_clean_sample <- df_TFT_clean_sample %>%
  filter(avg_unit_level != 0, board_cost != 0) 

plot(df_TFT_clean_sample[c(1, 7, 11, 12, 13)])
```

```{r}
df_TFT_clean_sample %>%
  ggplot() +
    geom_bar(aes(x=tier, fill=level), position='fill')

df_TFT_clean_sample %>%
  ggplot() +
    geom_bar(aes(x=tier, fill=placement), position='fill')

df_TFT_clean_sample %>%
  ggplot() +
    geom_boxplot(aes(x=tier, y=last_round))

df_TFT_clean_sample %>%
  ggplot() +
    geom_boxplot(aes(x=tier, y=avg_unit_level))

df_TFT_clean_sample %>%
  ggplot() +
    geom_boxplot(aes(x=tier, y=gold_left))

df_TFT_clean_sample %>%
  ggplot() +
    geom_boxplot(aes(x=tier, y=game_length))

df_TFT_clean_sample %>%
  ggplot() +
    geom_bar(aes(x=tier, fill=n_legend), position='fill')
```

After looking at all the plots above, I am not sure if game_length will be 
useful since the distribution is similar across tiers. Similarly, I will need to
keep my eye on n_legend, last_round, and gold_left.

```{r}
temp <- df_TFT_clean_sample %>%
  mutate(a1 = if_else(last_round >= 5, is.na(augment_1), 0),
         a2 = if_else(last_round >= 13, is.na(augment_2), 0),
         a3 = if_else(last_round >= 20, is.na(augment_3), 0))

temp %>%
  ggplot() +
  geom_bar(aes(x=tier, fill=as_factor(a1)), position='fill')

temp %>%
  ggplot() +
  geom_bar(aes(x=tier, fill=as_factor(a2)), position='fill')

temp %>%
  ggplot() +
  geom_bar(aes(x=tier, fill=as_factor(a3)), position='fill')
```


```{r}
write.csv(df_TFT_clean_sample, "TFT_data.csv", row.names = FALSE)
```

